---
title: "Final Project"
author: "Ni, Hugo"
date: "2024-02-29"
output: html_notebook
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Libaries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(shiny)
library(shinydashboard)
library(naniar)
library(janitor)
library(ggplot2)
```

# 1. Import the data
```{r}
broomrape <- read.csv("Table_2_Association%20mapping%20for%20broomrape%20resistance%20in%20sunflower.csv")
```

```{r}
#accession:Type of sunflower
#source: origin
#broomrape type_year. 2 per broomrape type measured in different year
#count: avg number of broomrapes on it.
```
# 2. Checks for NAs
```{r}
miss_var_summary(broomrape)
```

# 3.Clean the data column names
```{r}
broomrape<-
  clean_names((broomrape))
```

## remove unecessary columns
```{r}
broomrape <-
  broomrape %>%
  select(!confectionery)
```

## pivot data longer
```{r}
broomrape_long <- broomrape %>%
  pivot_longer(cols = c(3:8), names_to = "broomrape", values_to = "counts")
```

## separate broomrape type and year
```{r}
broomrape_sep<-
  broomrape_long %>%
  separate(broomrape, into=c("broomrape_name","year"),sep="_")
```

# 4. Find out which broomrape type is most parasitic
```{r}
broomrape_sep %>%
  group_by(broomrape_name) %>%
  summarise(average_count = mean(counts, na.rm = T))
```
## Plot construction
```{r}
broomrape_sep %>%
  group_by(broomrape_name) %>%
  summarise(average_count = mean(counts, na.rm = T)) %>%
  ggplot(aes(x=broomrape_name, y=average_count))+
  geom_col()+
  labs(title = "Average Attachment for Each Broomrape Type",
    x="Broomrape Names",
    y="Broomrape Counts")
```

# 5. Find out which host plant is most resistant
```{r}
broomrape_sep %>%
  group_by(accession) %>%
  summarise(ave_attachment = mean(counts, na.rm = T)) %>%
  arrange(ave_attachment)
```
## constructing plot
```{r}
broomrape_sep %>%
  group_by(accession) %>%
  summarise(ave_attachment = mean(counts, na.rm = T)) %>%
  arrange(ave_attachment) %>%
  top_n(-5) %>%
  ggplot(aes(x=accession , y=ave_attachment))+
  geom_col()+
  coord_flip()+
  labs(title = "Average Attachment in Different Sunflower",
    x="Sunflower Accession",
    y="Broomrape Counts")
```
# Looking at the distinct accessions in the data
```{r}
distinct(broomrape_long,accession)
```
# Shiny app. Our data was super diverse in terms of the number of different plants that had broomrapes on them. We want to compare the counts of different broomrapes based on what plant they were on.
```{r}
ui <- fluidPage(
  titlePanel("Broomrape Data Visualization"),
  sidebarLayout(
    sidebarPanel(
      selectInput("source", "Origin of host plant (where the host plant was collected from):",
                  choices = unique(broomrape_sep$source)),
      selectInput("accessions", "Select Accessions (Plant that hosted the broomrape):",
                  choices = NULL, multiple = TRUE)
    ),
    mainPanel(
      plotOutput("plot")
    )
  )
)


server <- function(input, output, session) {
  
  
  observe({
    updateSelectInput(inputId = "accessions",
                      choices = unique(broomrape_sep$accession[broomrape_sep$source == input$source]))
  })
  
 
  filtered_data <- reactive({
    broomrape_sep %>%
      filter(source == input$source) %>%
      filter(accession %in% input$accessions)
  })
  
  
  output$plot <- renderPlot({
    ggplot(filtered_data(), aes(x = accession, y = counts, fill = factor(year))) +
      geom_point(shape = 21, size = 4, position = position_dodge(width = 0.5)) +
      labs(title = "Broomrape Counts on Different Host Plants",
           x = "Host Plant", y = "Count of Broomrapes (on Avg)") +
      scale_fill_discrete(name = "Year") +  
      theme_bw() 
  })
}


shinyApp(ui = ui, server = server)
```




